---
title: "Activity2_Suicide"
output: html_document
date: "2024-08-06"
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(car)
library(MASS)
library(corrplot)
library(reshape2)
library(faraway)
library(caret)
library(nortest)
```

## 1. Import and preprocess data

```{r}
data <- read.csv("master.csv", header = TRUE)

attach(data)
head(data)
str(data)
dim(data)
```

### Rename columns
```{r}
data$gdp_for_year <- data$gdp_for_year....
data$gdp_for_year.... <- NULL

data$gdp_per_capita <- data$gdp_per_capita....
data$gdp_per_capita.... <- NULL
```

### Process missing values
```{r}
missing_values <- sapply(data, function(x) sum(is.na(x)))
print(missing_values)
```
The missing values of HDI.for.year account for about 70% of the dataset. Therefore, we decide to remove this variable. Moreover, country.year is combination string of country and year, which is redundant and suicides.100k.pop is suicides / (population / 100k), which includes our dependent variable suicides_no, so we also decide to remove those variables at first.

### Process unnecessary columns
```{r}
data_clean <- data
data_clean <- dplyr::select(data_clean, -HDI.for.year, -country.year, -suicides.100k.pop)
```

### Convert data types
```{r}
# Convert gdp_for_year to numeric after removing commas
data_clean <- data_clean %>%
mutate(gdp_for_year = as.numeric(gsub(",", "", gdp_for_year)))

str(data_clean)
detach(data)
attach(data_clean)
```
### Process duplicate rows
```{r}
duplicate_rows <- data_clean[duplicated(data_clean), ]
print(duplicate_rows)
```


### Descriptive statistic

```{r}
summary(data_clean)
```

```{r}
# Descriptive statistics for the 'suicides_no' column
hist(suicides_no, breaks = 30)
log_suicides_no <- log(suicides_no + 1) + 1

hist(log_suicides_no, breaks = 30)

boxplot(log_suicides_no)
```

```{r}
# Descriptive statistics for the 'year' column
boxplot(year)
hist(year, breaks = 20)
```
The most frequent years in the dataset are between 2000 and 2010, indicating that these years have the highest number of records related to suicides.

```{r}
# Descriptive statistics for 'sex' column
pie(table(sex))
barplot(table(sex))
```
The suicide records distributed evenly among male and female.





```{r}
# Descriptive Statistics for 'generation' and 'age'
barplot(table(generation))
pie(table(generation))
barplot(table(age))
pie(table(age))
```
Based on the observation, the suicides cases distributed evenly among different ages. The number of suicides records is the lowest in the Generation Z group.

```{r}
# Descriptive statistics for 'population'
boxplot(log(population))
hist(population, breaks = 30)
hist(log(population), breaks = 30)
hist(log(population)^2, breaks = 30)
```

```{r}
# Descriptive statistics for 'gdp_per_capita'
hist(gdp_per_capita, breaks = 50)
hist(log(gdp_per_capita), breaks = 50)
hist(log(gdp_per_capita)^2, breaks = 50)
```

```{r}
# Descriptive statistics for 'gdp_for_year'
boxplot(log(gdp_for_year))
hist(gdp_for_year, breaks = 30)
hist(log(gdp_for_year), breaks = 30)
hist(log(gdp_for_year)^2, breaks = 30)
```

### Preprocess data
```{r}
# Process 'suicides_no' column
data_clean$log_suicides_no <- log(suicides_no + 1) + 1
data_clean$suicides_no <- NULL
```


```{r}
# Process 'gdp_per_capita' column
data_clean$transform_gdp_per_capita <- log(gdp_per_capita)^2
data_clean$gdp_per_capita <- NULL
```

```{r}
# Process 'population' column
data_clean$transform_population <- log(population)^2
data_clean$population <- NULL
```


```{r}
# Process 'gdp_per_capita' column
data_clean$transform_gdp_for_year <- log(gdp_for_year)
data_clean$gdp_for_year <- NULL
```


```{r}
# Process 'year' column
data_clean$year <- data_clean$year - 1985 + 1
```


```{r}
# # Create dummy variables for the 'country' column

data_clean$regionEurope <- ifelse(data_clean$country %in% c("Austria", "Iceland", "Netherlands", "Belgium", "Bulgaria", "France", "Greece", "Ireland", "Italy", "Luxembourg", "Malta", "Norway", "Portugal", "Romania", "Spain", "Sweden", "United Kingdom", "Ukraine", "Finland", "Switzerland", "Serbia", "Slovenia", "Slovakia", "Albania", "Denmark", "Estonia", "Latvia", "Lithuania", "Belarus", "Croatia", "Czech Republic", "Germany", "Hungary", "Poland", "San Marino", "Bosnia and Herzegovina"), 1, 0)

data_clean$regionAsia <- ifelse(data_clean$country %in% c("Israel", "Japan", "Republic of Korea", "Singapore", "Turkmenistan", "Thailand", "Russian Federation", "Kazakhstan", "Kyrgyzstan", "Armenia", "Azerbaijan", "Philippines", "Cyprus", "Qatar", "Sri Lanka", "Maldives", "Turkey", "United Arab Emirates", "Oman", "Bahrain", "Uzbekistan", "Georgia", "Macau"), 1, 0)

data_clean$regionAfrica <- ifelse(data_clean$country %in% c("Mauritius", "South Africa", "Seychelles", "Cabo Verde"), 1, 0)

data_clean$regionNorthAmerica <- ifelse(data_clean$country %in% c("Canada", "Costa Rica", "Guatemala", "Mexico", "Puerto Rico", "United States", "Belize", "Saint Lucia", "Antigua and Barbuda", "Trinidad and Tobago", "Panama", "Saint Vincent and Grenadines", "Cuba", "El Salvador", "Bahamas", "Jamaica", "Saint Kitts and Nevis", "Dominica"), 1, 0)
                             
data_clean$regionSouthAmerica <- ifelse(data_clean$country %in% c("Argentina", "Brazil", "Chile", "Colombia", "Ecuador", "Paraguay", "Suriname", "Uruguay", "Guyana"), 1, 0)
                             
data_clean$regionOceania <- ifelse(data_clean$country %in% c("Australia", "New Zealand", "Fiji", "Kiribati", "Montenegro"), 1, 0)

# Remove the 'country' column after creating 'region'
data_clean$country <- NULL
```

```{r}
data_clean$sexMale <- ifelse(data_clean$sex == "male", 1, 0)
data_clean$sex <- NULL
```

```{r}
data_clean$geneX <- ifelse(data_clean$generation == "Generation X", 1, 0)
data_clean$geneMillenials <- ifelse(data_clean$generation == "Millenials", 1, 0)
data_clean$geneBoomers <- ifelse(data_clean$generation == "Boomers", 1, 0)
data_clean$geneSilent <- ifelse(data_clean$generation == "Silent", 1, 0)
data_clean$generation <- NULL
# The other generation would be G.I. Generation
```

```{r}
data_clean$Age15to24 <- ifelse(data_clean$age == "15-24 years", 1, 0)
data_clean$Age25to34 <- ifelse(data_clean$age == "25-34 years", 1, 0)
data_clean$Age35to54 <- ifelse(data_clean$age == "35-54 years", 1, 0)
data_clean$Age55to74 <- ifelse(data_clean$age == "55-74 years", 1, 0)
data_clean$age <- NULL
#The other generation would be 75+ years
```

```{r}
# Calculate Cook's distance
model <- lm(log_suicides_no ~ ., data = data_clean)
cooksD <- cooks.distance(model)

# Identify influential observations
influential <- cooksD[(cooksD > (3/length(cooksD)))]

# Extract the indices of influential observations
outliers <- as.numeric(names(influential))

# Remove influential observations to clean the data
data_clean <- data_clean[-outliers, ]
```

```{r}
plot(cooksD, type = "h", main = "Cook's Distance", ylab = "Cook's Distance")
abline(h = 4 / length(cooksD), col = "red") # original threshold line
abline(h = 3/ length(cooksD), col = "blue") # custom threshold line

```

### Analyze preprocessed dataset
```{r}
# Visualize preprocessed dataset
data_clean %>%
  gather(key = "variable", value = "value") %>%
  ggplot(aes(x = value)) +
  facet_wrap(~ variable, scales = "free") +
  geom_histogram(bins = 30) +
  theme_minimal()
data_clean %>%
  gather(key = "variable", value = "value") %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
# Descriptive statistics for the preprocessed dataset
summary(data_clean)
```


## 2. Split data to train and test
```{r}
set.seed(1)
sample_size <- floor(0.8 * nrow(data_clean))
train_indices <- sample(seq_len(nrow(data_clean)), size = sample_size)
data_train <- data_clean[train_indices, ]
data_test <- data_clean[-train_indices, ]
detach(data_clean)
attach(data_train)
```

## 3. Build model

### Check multicollinearity


```{r}
model = lm(log_suicides_no ~ ., data = data_train)
summary(model)
car::vif(model)
```

```{r}
model = lm(log_suicides_no ~ . -transform_gdp_for_year , data = data_train)
summary(model)
car::vif(model)
```

```{r}
cor(data_train$log_suicides_no, data_train$regionEurope)
model_regionEurope = lm(regionEurope ~ . -transform_gdp_for_year -log_suicides_no, data = data_train)
summary(model_regionEurope)
```


```{r}
model = lm(log_suicides_no ~ . -transform_gdp_for_year -regionEurope, data = data_train)
summary(model)
car::vif(model)
```

### Variable selection (Stepwise)

```{r}
modFull = lm(log_suicides_no ~ . -transform_gdp_for_year -regionEurope, data = data_train)
modZero = lm(log_suicides_no ~ 1, data = data_train)
modInter = lm(log_suicides_no ~ year + transform_population + regionAsia + regionNorthAmerica + sexMale + geneBoomers + Age25to34, data = data_train)
```

```{r}
modelAIC = MASS::stepAIC(modFull, direction = "backward", scope = list(lower = modZero, upper = modFull), k = 2)
```

```{r}
summary(modelAIC)
```

## 4. Hypothesis testing

### Durbin_Watson test for autocorrelation
```{r}
lmtest::dwtest(modelAIC)
```
### Shapiro-Wilk test for normality

```{r}
# Use Shapiro-Wilk test to test for normality of the residuals
set.seed(22) # For reproducibility
subsample_residuals <- sample(residuals(modelAIC), size = 5000)
shapiro.test(subsample_residuals)
```


```{r}
# Plot residuals to visually check for normality
par(mfrow=c(1,2))

hist_residuals <- residuals(modelAIC)
hist(hist_residuals, main = "Residuals Histogram", xlab = "Residuals", breaks = 30, probability = TRUE)

mean_residuals <- mean(hist_residuals)
sd_residuals <- sd(hist_residuals)
curve(dnorm(x, mean = mean_residuals, sd = sd_residuals), col = "red", add = TRUE)

qqnorm(hist_residuals, main = "Q-Q Plot of Residuals")
qqline(hist_residuals, col = "red")
```


### Studentized Breusch-Pagan test for heteroscedasticity
```{r}
bp_test <- lmtest::bptest(modelAIC)
print(bp_test)
```

### Box-Cox Transformation
```{r}
boxcox_result <- boxcox(modelAIC, plotit = TRUE)
lambda <- boxcox_result$x
log_likelihood <- boxcox_result$y

# Find the lambda with the maximum log-likelihood
best_lambda <- lambda[which.max(log_likelihood)]

# Print the best lambda
print(best_lambda)
```

```{r}
# Build the final model with Box-Cox transformation
best_lambda = 1.5
model_cox = lm((((data_train$log_suicides_no^best_lambda) - 1)/best_lambda) ~ 
    year + transform_gdp_per_capita + 
    transform_population + regionAsia + regionAfrica + regionNorthAmerica + 
    regionSouthAmerica + sexMale + geneX + geneMillenials + geneBoomers + 
    geneSilent + Age15to24 + Age25to34 + Age35to54 + Age55to74, data = data_train)
summary(model_cox)
```

```{r}
lmtest::dwtest(model_cox)
```


```{r}
# Use Shapiro-Wilk test to test for normality of the residuals
set.seed(22) # For reproducibility
subsample_residuals <- sample(residuals(model_cox), size = 5000)
shapiro.test(subsample_residuals)
```


```{r}
# Plot residuals to visually check for normality
par(mfrow=c(1,2))

hist_residuals <- residuals(model_cox)
hist(hist_residuals, main = "Residuals Histogram", xlab = "Residuals", breaks = 30, probability = TRUE)

mean_residuals <- mean(hist_residuals)
sd_residuals <- sd(hist_residuals)
curve(dnorm(x, mean = mean_residuals, sd = sd_residuals), col = "red", add = TRUE)

qqnorm(hist_residuals, main = "Q-Q Plot of Residuals")
qqline(hist_residuals, col = "red")
```


```{r}
# Studentized Breusch-Pagan test for heteroscedasticity
bp_test <- lmtest::bptest(model_cox)
print(bp_test)
```
The BP statistic has decreased from 3230.76 to 2943.80 after the Box-Cox transformation. This suggests that the transformation has helped reduce heteroscedasticity, but it has not completely eliminated it.

Despite the decrease in the BP statistic, the p-value remains extremely small, indicating that heteroscedasticity is still a significant issue. The transformation has not fully addressed the non-constant variance in the residuals.

## 6. Prediction

###Using cross validation k-folds
```{r}
set.seed(1)  
# Define a train control with k-fold cross-validation
train_control <- trainControl(method = "cv", number = 10)  # 10-fold cross-validation

# Define the formula for the model
formula <- as.formula(paste0("((log_suicides_no^", best_lambda, " - 1)/", best_lambda, ") ~  
    year + transform_gdp_per_capita + 
    transform_population + regionAsia + regionAfrica + regionNorthAmerica + 
    regionSouthAmerica + sexMale + geneX + geneMillenials + geneBoomers + 
    geneSilent + Age15to24 + Age25to34 + Age35to54 + Age55to74"))

# Train the model using the training data
cv_model <- train(formula, data = data_train, method = "lm", trControl = train_control)

```

```{r}
# Predict using the cross-validated model on the test data
predictions_log <- predict(cv_model, newdata = data_test)

# Calculate performance metrics on the test data
actual_values <- exp(data_test$log_suicides_no - 1) -1
predict_values <- round(exp((ifelse(predictions_log < 0, 0, predictions_log)  * best_lambda + 1)^(1/best_lambda) -1) -1, 2)
results <- data.frame(
  Actual = actual_values,
  Predicted = predict_values
)

# Print the results
print(results)
```

## 7. Evaluation
```{r}
# Calculate and print RMSE and R-squared
rmse <- sqrt(mean((results$Actual - results$Predicted)^2))
r_squared <- cor(results$Actual, results$Predicted)^2
cat("RMSE:", rmse, "\n")
cat("R-squared:", r_squared, "\n")
```